The truncated KdV framework for modeling anomalous waves induced by abrupt depth changes

### 2019-08-07 paper2_JNLS/paper.tex
NM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement; i.e.~When governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.

A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from variations in bottom topography. Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics. In particular, the more complex mechanism of focusing by 2D bottom topography absent. The studies thus offer the simplest conceivable system that exhibits the anomalous statistical behavior present in more complex systems, like the ocean, but in a tractable system that could be analyzed and quantified.
% ideal, paradigm, canonical, tractable

\vsp{4}
%Introduce waves. \\
Introduce papers: PRF, PNAS, and J Stat Phys. \\

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed both in deep and shallow water, and can be triggered by a variety of mechanisms, including anomalous wind-forcing, focusing due to variable bathymetry, and a deep-water modulational instability that goes by the name Benjamin-Feir. The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement. That is, when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these rare events can become much more common when surface statistics deviate from Gaussian.

A recent series of laboratory \cite{bolles2019anomalous} and numerical investigations \cite{viotti2014} have demonstrated the emergence of anomalous wave statistics from abrupt variations in bottom topography. Since topographical variations are strictly one-dimensional in these studies, they can be considered as offering a bare minimum set of conditions capable of generating anomalous wave statistics. In particular, the more complex mechanism of focusing by 2D bottom topography absent. The studies thus offer a paradigm system, with emergent anomalous statistical properties similar to that seen in more complex systems, but in a tractable context that is amenable to analysis.

Our particular focus is the laboratory experiments of Bolles et al.~\cite{bolles2019anomalous}, who demonstrated the emergence of anomalous statistics in a randomized wave-field that crosses an abrupt depth change (ADC). In these experiments, an incoming wave-field was generated with nearly Gaussian statistics. A plexiglass step created the abrupt depth change. Upon passing into a region of shallower depth, the wave distribution skewed strongly, with deviation from Gaussian being most pronounced a short distance downstream of the ADC. 
Inspired by these experimental observations, Majda et al.~\cite{majda2019statistical} laid out a theoretical foundation that accurately captures several aspects of the anomalous behavior. The theory is based on deterministic and statistical analysis of the truncated Korteweg–de Vries (TKdV) equations, and uses a combination of computational, statistical, and analytical tools. In follow-up work, Majda and Qi (cite) performed subsequent analysis on severely truncated, low-dimensional systems that were shown to exhibit many of the same anomalous features.

% ideal, paradigm, canonical, tractable
% ABRUPT depth change

\vsp{4}
Introduce papers: PRF, PNAS, and J Stat Phys. \\
